Prestige bias in cultural evolutionary dynamics

### 2023-11-16 ms/ms.tex
continue editing results

\begin{abstract}
If the traits of more successful individuals are more likely to be adopted, the resulting cultural transmission is described as being success-biased. In contrast, if the traits of `prestigious' individuals--those individuals that have already been copied many times--are more likely to be adopted, this is described as prestige-biased cultural transmission. In this case prestige can be used as a convenient proxy for success.
Here, we study the the joint effect of success and prestige bias on cultural evolutionary dynamics using mathematical analysis and stochastic simulations.
Analytic approximations to the stochastic role-model choice process facilitate the mathematical analysis and reduce the computational complexity of simulations. Approximations are given to the fixation probability and the fixation time of an invading cultural trait in different environments. 
We show that success bias effectively plays the role of natural selection, whereas prestige bias effectively plays the role of genetic drift.	Prestige bias, which may be strong in highly social communities, also accelerates the evolutionary dynamics, as can be expected in a rich-get-richer process.
\end{abstract}

@@

Here, we develop a stochastic model of cultural transmission with both indirect success bias and prestige bias to examine their relationship in contribution to the cultural evolution of populations.
We find analytic approximations for this model. We also find approximations for the probability and time to fixation of a `successful' phenotype (i.e., that is subject to success bias) in both a constant and a cyclic changing environment.
Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and prestige bias play the role of natural selection and genetic drift, respectively.

@@

Next, we examined the fixation probability and fixation time of a favored phenotype $\hat{A}$ when invading a population of phenotype $A$ and compared results from the full model and the DMD approximation.
Thus, we assume the population has a single individual with phenotype $\hat{A}$ and $N-1$ individuals with phenotype $A$. 
We find that the number of simulations needed to sufficiently approximate our model with the DMD approximation is roughly $1,000$ (\Cref{fig:num_sims}).
We examined the robustness of the DMD approximation to relaxing the following assumptions.
First, we relaxed our assumption of no estimation error $e$.
Estimation error in the original model was drawn from a normal distribution, and added to the trait value before evaluation of the bias ($A_{i,j} = A_j + e_i$).
When estimation error is applied, we sample $e_i$ for each copier $i$ from a normal distribution with expected value zero and variance $\eta^2$.
Even when the estimation error variance is $0.1$, both the fixation probability and fixation time approximations are accurate~(\Cref{fig:hetro_error}). 
We also relaxed our assumption of a uniform bias weight $\alpha$ (i.e., $\alpha_i=\alpha$). We allowed $\alpha$ to vary in the population, drawing $\alpha_j$ for each role-model $j$ from a normal distribution such that $\alpha_j \sim N(0.5,q)$ where $\epsilon$ is between $10^{-7}$ and $10^{-1}$. 
We found again that results of the DM approximation are similar to those from stochastic simulations of the full model~(\Cref{fig:hetro_alpha}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Fixation probability and time}
After finding that the DMD is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and conditional fixation time (conditioned on the population reaching fixation) of a favored phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the bias weight, $\alpha$, which determines the relative effect of success and prestige bias, given by \cref{eq:prestige}.

@@

\begin{result}[Effective selection coefficient and population size]\label{res:selection_coef}
The effective selection coefficient $s$ and effective population size $N_e$ can be written in terms of the success coefficient $\beta$ (\cref{eq:success_bias}), the bias weight $\alpha$ (\cref{eq:prestige}), and the population size $N$ as 
\begin{equation}
s=1-\beta=\frac{\beta(\hat{A})-\beta(A)}{\beta(\hat{A})}, \quad N_e=\alpha N + (1-\alpha) \;.
\end{equation}
\end{result}

@@

\paragraph{Numerical validation.}
We compare our approximations (\cref{eq:kimura_p,eq:kimura_t}) with results of simulations of our dichotomous model using various $\alpha$ and $\beta$ values, as well as simulations of the Wright-Fisher model, using the effective selection coefficient, $1-\beta$, and effective population size, $N_e=\alpha N + (1-\alpha)$. 
We find see that the two models have similar dynamics, and both are well approximated {by} our approximations (\Cref{fig:var_alpha}).

@@

\paragraph*{Changing environment}. After finding a good approximation in a constant environment, where the favorable trait is always $\hat{A}$, we proceeded to find an approximation for a cyclic changing environment. 
Following \citep{changeEnv}, we denote $k$ as the number of generations in which the invading phenotype is favored by success bias, and $l$ as the number of generations in which the resident phenotype is favored by success bias.
Thus, during the first $k$ generations of the environmental cycle, $\beta=\frac{\beta(A)}{\beta(\hat{A})}$<1, where $\hat{A}$ is the invading phenotype. During the following $l$ generations of the environmental cycle, the phenotype favored by success bias is switched, such that $\frac{\beta(A)}{\beta(\hat{A})}>1$.
We then proceed to find expressions for the expectation and variance of the change in the frequency of phenotype $\hat{A}$ after $n=k+l$ generations. 
The proof is in Appendix~\ref{sec:drift_diff_chang}.
\\

@@

Note that here, we have the average selection coefficient during a cycle of $n$ generations as the selection coefficient {in} \cref{eq:kimura_p}.
Using the drift and diffusion terms and following \citep{changeEnv}, we can approximate the fixation probability in a changing environment.
\\

\begin{result}[Fixation probability in a changing environment]\label{res:ch_expected}

@@

\paragraph{Numerical validation.}
To validate the the approximation for the fixation probability in a changing environment (\cref{eq:ch_env}), we compare it to results of simulations that use the DMD approximation (\Cref{cor:dirichlet}).
We find that the approximation fits the simulation results well for variable bias weights, $\alpha$, which corresponds to the effective population size (\Cref{fig:ch_env_alpha_beta}A).

However, the approximation is more sensitive to the value of the success bias coefficient $\beta$ (\Cref{fig:ch_env_alpha_beta}B).
We suspect that when $\beta$ is too small, there will not be many cycles in the simulations, because either the population reaches a high frequency of the fitter phenotype after just a few cycles, or the fitter phenotype {goes} extinct very quickly. 
For such $\beta$ values (below 0.65), the fixation probability exceeds even the constant environment approximation (which is the upper limit). We note that the diffusion-equation approximation assumes weak selection, or in our case, weak success bias.

We found that for a large $k/l$ ratio (with a constant cycle length, $n=k+l=100$), the changing environment approximation (\cref{eq:ch_env}) converges to the constant environment approximation (\cref{eq:kimura_p}), see \Cref{fig:ch_env_alpha_beta}C and \Cref{fig:ch_env_alpha_beta}D.

The approximation follows the trend of the simulation results for all $\alpha$ values.
{On} increasing the success coefficient {$\alpha$} to more than 0.15, the simulation results were located above the changing environment approximation, and below the constant environment approximation. We believe the reason is the structure of the cycle.
Our proof and approximation in the changing environment are for a large {number} of cycles, and when the success coefficient {$\alpha$} is too high, there might be very few cycles. Either the ideal trait is copied by enough copiers so that the prestige is sufficient to negate the success bias when the cycle changes (and the trait favored by the bias becomes the disfavored), or the opposite happens, and the ideal trait {goes} extinct before there {are} enough copiers that copied it.
We then tried to change the ratio between the number of cycles where $\hat{A}$ is favored and disfavored. We showed that the approximation fits well regardless of the ratio, but when the ratio of favored generations to disfavored ones is very high, it is very similar to a constant environment model.

\begin{figure}[H]
	\centering
    \includegraphics[width=\linewidth]{ch_env.pdf}
  \caption{\textbf{Fixation probability in a changing environment.}
\textbf{(A)} Fixation probability decreases with {the} success-bias weight (bottom x-axis) and effective population size (top x-axis). The approximation (blue; \cref{eq:ch_env}) agrees with simulation results (orange). 
\textbf{(B)} Fixation probability increases with the success coefficient, $\beta$.
When success bias is large ($1-\beta > 0.1$),  
simulation results (orange) are underestimated by the changing environment approximation (blue; \cref{eq:ch_env}). With even larger success bias ($1-\beta > 0.35$), even the constant environment approximation (green; \cref{eq:kimura_p}) slightly underestimates simulation results, likely because the diffusion equation approximation assumes weak "selection" .
(\textbf{C}) The approximation (blue) is robust to changes in environmental cycle length, as it agrees with simulations (orange) for different sizes of the changing environment cycle, where $k$ and $l$ are the number of generations each trait value is under success bias. 
When $k>l$, the approximation and the simulations are both very close to the constant environment approximation (green), because the more generations the rare phenotype is favored, the more similar it is to the constant environment model, where it is always favored by the success bias.
Markers show average of $10,000$ simulations, error bars show 75\% (panels A and C) and 95\% (panel B) confidence intervals.
Here, population size $N=1,000$; $\hat{A}=1$, $A=0.9$, $k=20$ and $l=80$ (panels A and B); $A=0.8$, (panel C); in panel A, the success coefficient is $1-\beta=s=0.005$; in panels B and C, the success-bias weight is $\alpha=0.1$.
  }
  \label{fig:ch_env_alpha_beta}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Optimal success-bias weight}
In results 2-6, we assumed that $\alpha$ is homogeneous in the population and constant, that is, it does not depend on any specific context.
Next, we examined what happens in the continuous-trait model if the $i$-th copier evaluates its own optimal success-bias weight, $\alpha^*_i$, which minimizes the expected squared error between the chosen trait value and the optimal trait value $\hat{A}$,
\begin{equation}\label{eq:optimal_alpha}
\alpha^*_i = \emph{argmin} \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_{i-1,j}}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_{i-1,l}} (\hat{A}-A_j)^2 \;,
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_{i-1,j}$ is the number of copiers that already chose role-model~$j$ by the time the $i$-the copier chooses a role-model.
Simply put, each copier $i$ estimates what success-bias weight $\alpha^*_i$ will result in copying a trait that is most similar to the ideal trait value $\hat{A}$.
Indeed, if the trait value is correlated with fitness, the optimal success-bias weight would increase the fitness of individuals. However, here we ignore the effects of natural selection, focusing instead on selectively neutral traits.

We find that when copiers choose their success-bias weight according to \cref{eq:optimal_alpha}, it quickly decreases with the number of copiers that have already chosen a role-model and then stays at what appears to be an equilibrium~(\Cref{fig:influence_advantage}).
Moreover, their estimation error is much lower compared to a constant success-bias weight, which gives roughly the same high estimation error to all copiers (compare \Cref{fig:influence_advantage}B and C): in this example, the optimal success-bias weight gives an estimation error (difference between the chosen and optimal trait) converges to $0.046$, whereas a constant success-bias weight gives values $>0.74$.

@@

We developed two cultural-evolutionary models, one with continuous and one with dichotomous trait values.
Our models include both indirect success and prestige biases, where the latter is a bias towards role-models with many copiers. We model these biases using a stochastic role-model choice process: each copier, in turn, randomly chooses a role-model, and this choice is affected both by the estimated success of each potential role-model and the number of copiers that already chose each role-model (eq.~\ref{eq:prestige}). 
Hence, our models have two ``nested'' stochastic processes: the role-model choice process within each generation, and the cultural-evolutionary process between generations.
To simplify the mathematical and computational analysis, we developed analytic approximations for the role-model choice process using the the {\em generalized binomial distribution} (GBD, \Cref{res:GBD}) and the {\em Dirichlet-Multinomial distribution} (DMD, \Cref{cor:dirichlet}).
The latter is especially useful, as it approximates the entire role-model choice process and only requires us to assume that the relative effect of success and prestige is a characteristic of the role-model and not the copier.

Analyzing the dichotomous-trait model using the the DMD approximation, we found approximations for the fixation probability and fixation time of a cultural trait under biased transmission in a constant environment.
Our approximations are similar to Kimura's evolutionary-genetic approximations, in that (i) the strength of success bias towards the invading cultural trait, $1-\beta(A)$, is equivalent to the selection coefficient in favor of a beneficial allele, $s$, and (ii) decreasing the relative weight of success versus prestige bias, $\alpha$, decreases the effective population size, $N_e$~(\Cref{fig:var_alpha}).

We also analyzed the dichotomous--trait model in a cyclic changing environment in which the identity of the success-biased trait switches after a fixed amount of generations~(\Cref{fig:ch_env_alpha_beta}).
We find that, similarly to the constant environment approximation, a change in the success-bias weight $\alpha$ has no negative effects on the goodness-of-fit of the approximation to simulation results.
We also showed that this approximation is more sensitive to changes in the success coefficient $\beta$ than the constant environment approximation, and a lower value is required to have a good fit. The ratio between the number of {generations} in which the rare phenotype is under positive transmission bias and the number of generations in which it is under negative bias does not affect the goodness-of-fit of the approximation.

>>>>

\begin{abstract}
If the traits of more successful individuals are more likely to be adopted, the resulting cultural transmission is described as being success-biased. In contrast, if the traits of `prestigious' individuals--those individuals that have already been copied many times--are more likely to be adopted, this is described as prestige-biased cultural transmission. In this case prestige can be used as a convenient proxy for success.
Here, we study the joint effect of success and prestige bias on cultural evolutionary dynamics using mathematical analysis and stochastic simulations.
Analytic approximations to the stochastic role-model choice process facilitate the mathematical analysis and reduce the computational complexity of simulations. Approximations are given to the fixation probability and the fixation time of an invading cultural trait in different environments. 
We show that success bias effectively plays the role of natural selection, whereas prestige bias effectively plays the role of genetic drift.	Prestige bias, which may be strong in highly social communities, also accelerates the evolutionary dynamics, as can be expected in a rich-get-richer process.
\end{abstract}

@@

Here, we develop a stochastic model of cultural transmission with both indirect success bias and prestige bias to examine their relationship in contribution to the cultural evolution of populations.
We find analytic approximations for this model. We also find approximations for the probability and time to fixation of a `successful' phenotype (i.e., that is subject to success bias) in both a constant and a periodically changing environment.
Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and prestige bias play the role of natural selection and genetic drift, respectively.

@@

Next, we examined the fixation probability and fixation time of a `successful' phenotype $\hat{A}$ (i.e., that is favored by success bias) when invading a population of phenotype $A$ and compared results from the full model and the DMD approximation.
Thus, we assume the population has a single individual with phenotype $\hat{A}$ and $N-1$ individuals with phenotype $A$. 
We find that the number of simulations needed to sufficiently approximate our model with the DMD approximation is roughly $1,000$ (\Cref{fig:num_sims}).
We examined the robustness of the DMD approximation to relaxing the following assumptions.
First, we relaxed our assumption of no estimation error $e$.
Estimation error in the original model was drawn from a normal distribution, and added to the trait value before evaluation of the success `bias ($A_{i,j} = A_j + e_i$).
When estimation error is applied, we sample $e_i$ for each copier $i$ from a normal distribution with expected value zero and variance $\eta^2$.
Even when the estimation error variance is $0.1$, both the fixation probability and fixation time DMD approximations are accurate~(\Cref{fig:hetro_error}). 
We also relaxed our assumption of a uniform bias weight $\alpha$ (i.e., $\alpha_i=\alpha$). We allowed $\alpha$ to vary in the population, drawing $\alpha_j$ for each role-model $j$ from a normal distribution such that $\alpha_j \sim N(0.5, \epsilon)$ where $\epsilon$ is between $10^{-7}$ and $10^{-1}$. 
We found again that results of the DMD approximation are similar to those from stochastic simulations of the full model~(\Cref{fig:hetro_alpha}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Fixation probability and time}
After finding that the DMD is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and conditional fixation time (conditioned on the population reaching fixation) of a `successful' phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the success-bias weight, $\alpha$, which determines the relative effect of success and prestige bias, given by \cref{eq:prestige}.

@@

\begin{result}[Effective selection coefficient and population size]\label{res:selection_coef}
The effective selection coefficient $s$ and effective population size $N_e$ can be written in terms of the success coefficient $\beta$ (\cref{eq:success_bias}), the success-bias weight $\alpha$ (\cref{eq:prestige}), and the population size $N$ as 
\begin{equation}
s=1-\beta=\frac{\beta(\hat{A})-\beta(A)}{\beta(\hat{A})}, \quad N_e=\alpha N + (1-\alpha) \;.
\end{equation}
\end{result}

@@

\paragraph{Numerical validation.}
We compare our approximations (\cref{eq:kimura_p,eq:kimura_t}) with results of simulations of our dichotomous model using various $\alpha$ and $\beta$ values, as well as simulations of the Wright-Fisher model, using the effective selection coefficient, $1-\beta$, and effective population size, $N_e=\alpha N + (1-\alpha)$. 
We find that the two models have similar dynamics, and both are well approximated {by} our approximations (\Cref{fig:var_alpha}).

@@

\paragraph*{Changing environment}. After finding a good approximation in a constant environment, where the `successful' trait is always $\hat{A}$, we proceeded to find an approximation for a periodically changing environment. 
Following Ram et al.~\citep{changeEnv}, we denote $k$ as the number of generations in which the invading phenotype is favored by success bias, and $l$ as the number of generations in which the resident phenotype is favored by success bias.
Thus, during the first $k$ generations of the environmental cycle, $\beta=\frac{\beta(A)}{\beta(\hat{A})}$<1, where $\hat{A}$ is the invading phenotype. During the following $l$ generations of the environmental cycle, the phenotype favored by success bias is switched, such that $\frac{\beta(A)}{\beta(\hat{A})}>1$.
We then proceed to find expressions for the expectation and variance of the change in the frequency of phenotype $\hat{A}$ after $n=k+l$ generations. 
The proof is in Appendix~\ref{sec:drift_diff_chang}.
\\

@@

Note that here, we have the `average selection coefficient' during a cycle of $n$ generations, $S_n/n$ as the selection coefficient {in} \cref{eq:kimura_p}.
Using the drift and diffusion terms and following Ram et al.~\citep{changeEnv}, we can approximate the fixation probability in a changing environment.
\\

\begin{result}[Fixation probability in a periodically changing environment]\label{res:ch_expected}

@@

Importantly, the average selection coefficient, $S_n/n$, has the same sign as $k-l$. Therefore, when $k>l$, the fixation probability will increase with the success-bias weight $\alpha$ (similar to a constant environment, \Cref{fig:var_alpha}A), and when $k<l$, the fixation probability will decrease with the success-bias weight $\alpha$ (\Cref{fig:ch_env_alpha_beta}A).
Similarly, when the success coefficient $1-\beta$ is large, the approximation can break, as the fixation time can be lower then the number of generations in the cycle, $n$ (see \Cref{fig:var_alpha}D), and again the average selection coefficient, $S_n/n$ is not a good estimate of the effective selection coefficient.

\paragraph{Numerical validation.}
To validate the approximation for the fixation probability in a changing environment (\cref{eq:ch_env}), we compare it to results of simulations that use the DMD approximation (\Cref{cor:dirichlet}).
We find that the approximation fits the simulation results well for variable success-bias weights, $\alpha$, which corresponds to the effective population size (\Cref{fig:ch_env_alpha_beta}A).
However, the approximation is more sensitive to the value of the success bias coefficient $\beta$ (\Cref{fig:ch_env_alpha_beta}B).
On increasing the success coefficient $1-\beta$ to more than 0.15, the simulation results were located above the changing environment approximation, and below the constant environment approximation (\Cref{fig:ch_env_alpha_beta}B). This is probably because when the success coefficient is high, the number of generations may be less than $n$, the number of generations in a full cycle, and thus $S_n/n$ is not a good estimator for the `average selection coefficient'.
With even higher success coefficient values, the fixation probability exceeds even the constant environment approximation, possibly because the diffusion-equation approximation assumes weak selection, or in our case, weak success bias.
We also changed the ratio between the number of cycles where $\hat{A}$ is favored and disfavored. We found that the approximation fits well regardless of the ratio, and that for a large $k-l$ difference (with a constant cycle length, $n=k+l=100$), the changing environment approximation (\cref{eq:ch_env}) converges to the constant environment approximation (\cref{eq:kimura_p}; \Cref{fig:ch_env_alpha_beta}C). This makes sense as a constant environment can be viewed as an environment in which the cycle length is longer than the fixation time.

\begin{figure}[H]
	\centering
    \includegraphics[width=\linewidth]{ch_env.pdf}
  \caption{\textbf{Fixation probability in a changing environment.}
\textbf{(A)} Fixation probability decreases with {the} success-bias weight (bottom x-axis) and effective population size (top x-axis). The approximation (blue; \cref{eq:ch_env}) agrees with simulation results (orange). 
\textbf{(B)} Fixation probability increases with the success coefficient, $\beta$.
When success bias is large ($1-\beta > 0.1$),  
simulation results (orange) are underestimated by the changing environment approximation (blue; \cref{eq:ch_env}). With even larger success bias ($1-\beta > 0.35$), even the constant environment approximation (green; \cref{eq:kimura_p}) slightly underestimates simulation results, likely because the diffusion equation approximation assumes weak `selection'.
(\textbf{C}) The approximation (blue) is robust to changes in environmental cycle length, as it agrees with simulations (orange) for different sizes of the changing environment cycle, where $k$ and $l$ are the number of generations each trait value is under success bias. 
When $k>l$, the approximation and the simulations are both very close to the constant environment approximation (green), because the more generations the rare phenotype is favored, the more similar it is to the constant environment model, where it is always favored by the success bias.
Markers show average of $10,000$ simulations, error bars show 75\% (panels A and C) and 95\% (panel B) confidence intervals.
Here, population size $N=1,000$; $\hat{A}=1$, $A=0.9$, $k=20$ and $l=80$ (panels A and B); $A=0.8$, (panel C); in panel A, the success coefficient is $1-\beta=s=0.005$; in panels B and C, the success-bias weight is $\alpha=0.1$.
  }
  \label{fig:ch_env_alpha_beta}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Optimal success-bias weight}
In results 2-6, we assumed that $\alpha$ is homogeneous in the population and constant, that is, it does not depend on any specific context.
Next, we examined what happens in the continuous-trait model if the $i$-th copier evaluates its own optimal success-bias weight, $\alpha^*_i$, which minimizes the expected squared error between the chosen trait value and the `successful' trait value $\hat{A}$,
\begin{equation}\label{eq:optimal_alpha}
\alpha^*_i = \emph{argmin} \sum_{j=1}^N\frac{\alpha A_j + (1-\alpha) K_{i-1,j}}{\sum_{l=1}^N\alpha A_l + (1-\alpha) K_{i-1,l}} (\hat{A}-A_j)^2 \;,
\end{equation}
where $A_j$ is the trait of role-model $j$ and $K_{i-1,j}$ is the number of copiers that already chose role-model~$j$ by the time the $i$-the copier chooses a role-model.
Simply put, each copier $i$ estimates what success-bias weight $\alpha^*_i$ will result in copying a trait that is most similar to the `successful' trait value $\hat{A}$.
Indeed, if the trait value is correlated with fitness, the optimal success-bias weight would increase the fitness of individuals. However, here we ignore the effects of natural selection, focusing instead on selectively neutral traits.

We find that when copiers choose their success-bias weight according to \cref{eq:optimal_alpha}, it quickly decreases with the number of copiers that have already chosen a role-model and then stays at what appears to be an equilibrium~(\Cref{fig:influence_advantage}).
Moreover, the estimation error is much lower compared to a constant success-bias weight, which gives roughly the same high estimation error to all copiers (compare \Cref{fig:influence_advantage}B and C): in this example, the optimal success-bias weight gives an estimation error (difference between the chosen and `successful' trait) that converges to $0.046$, whereas a constant success-bias weight gives values greater than 0.74.

@@

We developed two cultural-evolutionary models, one with continuous and one with dichotomous trait values.
Our models include both indirect success and prestige biases, where the latter is a bias towards role-models with many copiers. We model these biases using a stochastic role-model choice process: each copier, in turn, randomly chooses a role-model, and this choice is affected both by the estimated success of each potential role-model and the number of copiers that already chose each role-model (eq.~\ref{eq:prestige}). 
Hence, our models have two ``nested'' stochastic processes: the role-model choice process within each generation, and the cultural-evolutionary process between generations.
To simplify the mathematical and computational analysis, we developed analytic approximations for the role-model choice process using the {\em generalized binomial distribution} (GBD, \Cref{res:GBD}) and the {\em Dirichlet-Multinomial distribution} (DMD, \Cref{cor:dirichlet}).
The latter is especially useful, as it approximates the entire role-model choice process and only requires us to assume that the relative effect of success and prestige is a characteristic of the role-model and not the copier.

Analyzing the dichotomous-trait model using the DMD approximation, we found approximations for the fixation probability and fixation time of a cultural trait under biased transmission in a constant environment.
Our approximations are similar to Kimura's evolutionary-genetic approximations, in that (i) the strength of success bias towards the invading cultural trait, $1-\beta(A)$, is equivalent to the selection coefficient in favor of a beneficial allele, $s$, and (ii) decreasing the relative weight of success versus prestige bias, $\alpha$, decreases the effective population size, $N_e$~(\Cref{fig:var_alpha}).

We also analyzed the dichotomous--trait model in a periodically changing environment in which the identity of the success-biased trait switches after a fixed amount of generations~(\Cref{fig:ch_env_alpha_beta}).
We find that, similarly to the constant environment approximation, a change in the success-bias weight $\alpha$ has no negative effects on the goodness-of-fit of the approximation to simulation results.
We also showed that this approximation is more sensitive to changes in the success coefficient $\beta$ than the constant environment approximation, and a lower value is required to have a good fit. The ratio between the number of {generations} in which the rare phenotype is under positive transmission bias and the number of generations in which it is under negative bias does not affect the goodness-of-fit of the approximation.
